\section*{Appendix B --- Glossary}\label{sec:appendix-b}
\addcontentsline{toc}{section}{Appendix B --- Glossary}
\begin{description}
  \item[Holonomy] Loop-induced phase from parallel transport around the fiber:
  \[
    \Delta\phi_\theta=\frac{q_\theta}{\hbar}\oint A_\theta\,d\theta \equiv \phi_\theta\ (\bmod\ 2\pi).
  \]
  \begin{itemize}
    \item \textbf{Analogy:} hiking a loop around a hill and finding your compass rotated when you return.
    \item \textbf{Where:} \cref{sec:theta-ab}.
  \end{itemize}

  \item[Mixed curvature $G_{\mu\theta}$] Coupling between base and fiber:
  \[
    G_{\mu\theta}=\partial_\mu A_\theta-\partial_\theta A_\mu,\quad \text{with}\; \partial_\theta A_\mu=0\Rightarrow G_{i\theta}=\partial_i A_\theta.
  \]
  \begin{itemize}
    \item \textbf{Analogy:} a gear train linking forward motion and an internal wheel.
    \item \textbf{Where:} \cref{sec:cross-hall}.
  \end{itemize}

  \item[Minisuperspace] Truncated configuration space for homogeneous modes (e.g., $(a,\theta)$ in FRW).
  \begin{itemize}
    \item \textbf{Analogy:} a city map showing only two main avenues.
    \item \textbf{Where:} \cref{sec:cosmology}.
  \end{itemize}

  \item[Inverse-square barrier] Repulsive $+C/a^2$ term in the Wheeler--DeWitt (WDW) equation, with threshold for essential self-adjointness at $\gamma\ge 3/4$ for $-\chi''+\tfrac{\gamma}{a^2}\chi$.
  Here $C\propto \ell^2\hbar^2/I_0$.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:wdw}.
  \end{itemize}

  \item[Phase stiffness / inertia $I$] Sets rotor sideband spacing $\Delta E\approx \hbar^2/(2I)$. NR map: $I=m\kappa^2$.
  \begin{itemize}
    \item \textbf{Analogy:} a heavier flywheel has more closely spaced levels.
    \item \textbf{Where:} \cref{sec:sidebands}.
  \end{itemize}

  \item[Einbein $e(\tau)$] Worldline multiplier enforcing reparametrization invariance; varying it imposes the mass-shell constraint. In the relativistic map one finds $I=m\kappa^2$.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:relativistic-worldline}.
  \end{itemize}

  \item[Gauge connection $A$] One-form on $Q$ that tracks phase transport:
  $A= A_\mu\,dX^\mu + A_\theta\,d\theta$, with curvature $G=dA$.
  \begin{itemize}
    \item \textbf{Analogy:} navigating a boat with changing currents.
    \item \textbf{Where:} \cref{sec:gauge-invariance}.
  \end{itemize}

  \item[Compact dimension $S^1$] Fiber coordinate $\theta$ is an angle with period $2\pi$.
  \begin{itemize}
    \item \textbf{Analogy:} a garden hose looks 1D from far away but has a circular cross-section up close.
    \item \textbf{Where:} \cref{sec:framework}.
  \end{itemize}

  \item[Flux quanta and large gauge] The $U(1)_\theta$ flux quantum is $\Phi_{\theta,0}=2\pi\hbar/q_\theta$; large $\theta$-loops shift $\oint A_\theta d\theta$ by $2\pi\hbar/q_\theta$, so only $\phi_\theta$ modulo $2\pi$ is physical.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:gauge-invariance}.
  \end{itemize}

  \item[Rotor (internal) and sidebands] With $\psi\sim e^{i\ell\theta}$, levels are
  \[
    E_\ell=\frac{\hbar^2}{2I}\Big(\ell-\frac{\phi_\theta}{2\pi}\Big)^2,\quad \ell\in\mathbb Z,
  \]
  giving near-harmonic sidebands spaced by $\Delta E\approx \hbar^2/(2I)$.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:sidebands}.
  \end{itemize}

  \item[Stueckelberg mass $m_\theta$ and Yukawa range $\lambda_\theta$] In 4D, $m_\theta=g_\theta f_\theta$ and $\lambda_\theta\equiv 1/m_\theta$. The same $\lambda_\theta$ controls short-range fingerprints across gravity/QED/weak sectors.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:unified-force}.
  \end{itemize}

  \item[Kinetic mixing $\varepsilon$] $U(1)$--$U(1)$ mixing induces a small Yukawa bump in Coulomb's law.
  \begin{itemize}
    \item \textbf{Analogy:} two pendulums tied by a weak spring.
    \item \textbf{Where:} \cref{sec:unified-force}.
  \end{itemize}

  \item[Cross-Hall drift] Transverse drift proportional to $\partial_i A_\theta$ when the fiber potential varies across space; see also Appendix~\ref{sec:appendix-a} for a paraxial coefficient.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:cross-hall}.
  \end{itemize}

  \item[Charges $q_X,\ q_\theta$] Minimal coupling rules: $\mathbf p\to\mathbf p-q_X\mathbf A$ and $p_\theta\to p_\theta-q_\theta A_\theta$.
  \begin{itemize}
    \item \textbf{Where:} \cref{sec:framework}.
  \end{itemize}
\end{description}
